---
title: "Gas Mileage Analysis of Motor Trend Data"
author: "David Maust"
date: "06/21/2014"
output: pdf_document
---

# Executive Summary

When selecting a car, fuel efficiency is a common selection criteria. In this paper, we look at 1974 Motor Trend data [1] for the purpose of evaluating factors on fuel efficiency. We specifically are interested in the effects of automatic vs manual transmission on the gas mileage.

# Preprocessing the Data

We begin by loading the data, and casting the transmission type and number of cylinders to factors.

```{r}
data(mtcars)
mtcars$am <- factor(mtcars$am, levels=c(0,1), labels=c('Automatic', 'Manual'))
mtcars$cyl <- as.factor(mtcars$cyl)
```

# Model Selection

We can use the step function found in the stats package of R to perform model selection.
```{r results='hide'}
model <- lm(mpg ~ am + wt + cyl + disp + hp + carb + gear + vs + drat, mtcars)
final_model <- step(model, direction='backward')
```

The final model returned only depends on the transmission type, number of carburetors, and displacement. Other variables did not contribute significantly to the gas mileage usage of the vehicle.

```{r}
summary(final_model)
```
In this, the final model has the transmission type, number of carburetors, and displacement. In it we can see that when we hold displacement, and the number of carburetors constant, a manual transmission will achieve `r coef(final_model)['amManual']` better mpg.

Looking at a residual plot, we see no patterns.
```{r}
plot(resid(final_model) ~ predict(final_model))
```

Looking at a QQ plot of residual we can see the residuals are approximately normal.
```{r}
qqnorm(resid(final_model))
qqline(resid(final_model))
```

# References

[1] Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.

# Appendix

## Figure 1: mpg comparision with respect to weight.
```{r}
plot(mpg ~ wt, mtcars, col=am, xlab = 'Displacement')
legend('topright', lty=c(1,1), col=1:2, legend = c('Automatic', 'Manual'))
```

## Figure 2: mpg comparision with respect to number of cylinders.
```{r}
plot(mpg ~ cyl, mtcars, col=am)
legend('topright', lty=c(1,1), col=1:2, legend = c('Automatic', 'Manual'))
```

## Figure 3: mpg comparision with respect to horsepower.
```{r}
plot(mpg ~ hp, mtcars, col=am)
legend('topright', lty=c(1,1), col=1:2, legend = c('Automatic', 'Manual'))
```

