---
title: "Gas Milage Analysis of Motor Trend Data"
author: "David Maust"
date: "06/21/2014"
output: pdf_document
---

# Executive Summary

When selecting a car, fuel efficiency is a common selection criteria. In this paper, we look at 1974 Motor Trend data [1] for the purpose of evaluating factors on fuel efficiency.

# Preprocessing the Data

```{r}
data(mtcars)
mtcars$am <- factor(mtcars$am, levels=c(0,1), labels=c('Automatic', 'Manual'))
```

# Model Selection
```{r results='hide'}
model <- lm(mpg ~ am + cyl + disp + hp + carb + gear + vs + qsec + drat, mtcars)
final_model <- step(model)
```

## Final Model
```{r}
summary(final_model)
```
In this, the final model has the transmission type, number of carburetors, and displacement. In it we can see that when we hold displacement, and the number of carburetors constant, a manual transmission will achieve `r coef(final_model)[2]` better mpg.

# References

[1] Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.

# Appendix

## Figure 1: mpg comparision with respect to displacement.
```{r}
library(ggplot2)
ggplot(mtcars, aes(x=disp, y=mpg, group=am, color=am)) + geom_point()
```

## Figure 2: mpg comparision with respect to number of carburetors.
```{r}
ggplot(mtcars, aes(x=carb, y=mpg, group=am, color=am)) + geom_point()
```

